#title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true



## Loading and preprocessing the data
```{r loadData, results="hide"}
library(dplyr)
rawactivity <- read.csv("activity.csv", header = TRUE)
rawactivity$date <- as.Date(as.character(rawactivity$date), "%Y-%m-%d")

```


## What is mean total number of steps taken per day?
```{r totalSteps}
activity <- rawactivity[!is.na(rawactivity$steps),]
aggactivity <- aggregate(activity$steps, list(activity$date), sum)
meanStep <- mean(aggactivity$x)
medianStep <- median(aggactivity$x)
meanStep
medianStep
hist(aggactivity$x, breaks=10, xlab = 'Steps')
```
The average totoal number of steps is `r meanStep`, and the median total number of steps is `r medianStep`.

## What is the average daily activity pattern?
```{r stepsInterval}
intervalGroups <- group_by(activity, interval)
intervalReport <- summarize(intervalGroups, averageSteps=mean(steps))
plot(intervalReport$interval, intervalReport$averageSteps, type = "l" )
maxInterval <- intervalReport[(intervalReport$averageSteps == max(intervalReport$averageSteps)), c("interval")]

```
The interval of `r maxInterval` minutes, on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r fillMissingValues}
naIndexes <- which( is.na(rawactivity$steps))
totalMissingValues <- length(naIndexes)
fullactivity <- rawactivity

for(index in naIndexes){
  inter <- fullactivity[index,]$interval
  avgIndex <- which(intervalReport$interval == inter)
  fullactivity[index,]$steps <- intervalReport[avgIndex, ]$averageSteps
}

aggactivity <- aggregate(fullactivity$steps, list(fullactivity$date), sum)
filledmeanStep <- mean(aggactivity$x)
filledmedianStep <- median(aggactivity$x)
filledmeanStep
filledmedianStep
hist(aggactivity$x, breaks=10, xlab = 'Steps')

```
The total number of missing values in the dataset is `r totalMissingValues`.  In order to fill the missing data, use the average steps on the same time in other days.  For example, if the step is NA at interval 25 on 2012-10-02, use the average step value of inteval 25 from other days.

The average step filled data of averages is `r filledmeanStep`, and the median step with filled data of average is `r filledmedianStep`.  The values are the same as the filled values are just average from the real data on the same time of the day.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdayPattern}
activity$weekday <- "weekday"
weekendIndex <-  which(format(activity$date, "%a") %in% c("Sat", "Sun"))
activity[weekendIndex,]$weekday <- "weekend"
activity$weekday <- factor(activity$weekday)
weekGroups <- group_by(activity, interval, weekday)
weekReport <- summarize(weekGroups, averageSteps=mean(steps))

library(lattice)
xyplot(averageSteps ~ interval | weekday , data =weekReport, type = "l" , layout=c(1,2), ylab="Number of Steps")

```
Based on the Plot, the pattern between weekend and weekend are similar.

